---
title: "Tidy-02-Join-Daily"
author: "David Barkemeyer"
date: "2023-03-12"
output: html_document
params:
  options: "./options.yaml"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Join of Trends and Daily CRSP

```{r}
library(lubridate)
library(readr)
library(zoo)
library(yaml)

options <- read_yaml(params$options)

crsp_daily <- read_rds("./../large_data/crsp_daily.rds")
trends <- read_rds(
  file.path("./../large_data", paste_time0(options), "trends.rds")
)

df <- full_join(trends, crsp_daily, by = c("date", "ticker"))

names(df)
```

# Arrange by ticker and date

```{r}
df_2 <- df %>%
  arrange(ticker, date)
```

# Remove ticker for which no CRSP data exists

```{r}
df_3 <- df_2 %>%
  group_by(ticker) %>%
  mutate(no_crsp = all(is.na(ret))) %>%
  ungroup() %>%
  filter(!no_crsp)
```

# Replace NAs in (nearly) static columns with previous observations

These (nearly) static columns are from the CRSP data and are therefore NA on
non-trading days. However, because they are (nearly) static it is safe to
replace them with their previous non-NA value.

```{r}
df_4 <- df_3 %>%
  group_by(ticker) %>%
  mutate(
    across(
      c(start_date, end_date, exchange, industry),
      ~ na.locf(na.locf(., na.rm = F), fromLast = T)
    )
  ) %>%
  ungroup()
```

# Remove observations before start date or after end date

```{r}
df_5 <- df_4 %>%
  filter(date >= start_date, date <= end_date)
```

# Market Capitalization (Size)

```{r}
df_6 <- df_5 %>%
  mutate(
    market_cap = prc * shares_outstanding
  )
```

# Save Data

```{r}
write_rds(
  df_6, 
  file.path("./../large_data", paste_time0(options), "crsp_trends.rds")
)
```

