---
title: "AbnormalReturns"
author: "David Barkemeyer"
date: "2023-03-13"
output: html_document
params: 
  options: "./options.yaml"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r}
library(lubridate)
library(readr)
library(yaml)
```

# Load data

```{r}
compustat <- read_rds("./../large_data/compustat.rds")

options <- read_yaml(params$options)

crsp_trends <- read_rds(
  file.path("./../large_data", paste_time0(options), "crsp_trends.rds")
)

crsp_monthly <- read_rds("./../large_data/crsp_monthly.rds")
```

# Select relevant columns

```{r}
crsp_trends <- crsp_trends %>%
  select(
    date, dly_attention, ticker, exchange, prc, shares_outstanding, ret, volume,
    market_cap
  )
```

# Size as of last trading day in June

```{r}
size <- crsp_trends %>%
  filter(
    month(date) == 6,
    !is.na(ret)
  ) %>%
  mutate(period = year(date)) %>%
  group_by(ticker, period) %>%
  filter(day(date) == max(day(date))) %>%
  mutate(
    period = year(date),
    size = prc * shares_outstanding
  ) %>%
  select(ticker, period, size) %>%
  ungroup()
```

# Book-to-market

```{r}
book_to_market <- compustat %>%
  mutate(period = year(date) + 1) %>%
  select(ticker, period, book_to_market)
```


# Proxy momentum through annual return of the previous twelve months

Annual return is computed from monthly returns from June of the previous year to May of the current year.

```{r}
momentum_period <- function(y, m) {
  # month greater than or equal to June -> count towards next year
  # month less than or equal to May -> count towards current year
  ifelse(m >= 6, y + 1, y)
}

momentum <- crsp_monthly %>%
  mutate(period = momentum_period(year(date), month(date))) %>%
  group_by(ticker, period) %>%
  summarise(momentum = average_return(ret)) %>% 
  suppressMessages() %>%
  ungroup()
```

# Join the three characteristics data sets

```{r}
factors <- inner_join(size, book_to_market, by = c("ticker", "period")) %>%
  inner_join(momentum, by = c("ticker", "period"))
```

# Join the factors with CRSP & trends data

```{r}
portfolio_period <- function(date) {
  # month greater than or equal to July -> count towards this year
  # month less than or equal to June -> count towards last year
  ifelse(month(date) >= 7, year(date), year(date) - 1)
}

# Prepare crsp_trends for join
ct <- crsp_trends %>%
  mutate(period = portfolio_period(date)) %>%
  select(-prc, -shares_outstanding)

data <- inner_join(ct, factors, by = c("ticker", "period"))
```

# For each period form quintiles on size, book-to-market and momentum (dependent sort)

## Portfolios on size

To follow the approach of Daniel et al. (1997) the size quintiles are computed based on NYSE stocks.

```{r}
probs <- seq(0, 1, by = 0.2) # quintile probabilities

# Compute size quintiles per twelve-month period
size_quintiles <- data %>%
  filter(exchange == "N") %>%
  select(date, period, ticker, size) %>%
  group_by(period) %>%
  summarise(size_5 = list(quantile(size, probs = probs)))

# Join size quintiles to ticker-period-size data
data_2 <- data %>%
  select(ticker, period, size, book_to_market, momentum) %>%
  distinct() %>%
  inner_join(size_quintiles, by = "period")

# Compute bin in which ticker-size combination is put per period
data_3 <- data_2 %>%
  rowwise() %>%
  mutate(size_q = findInterval(size, size_5, all.inside = T)) %>%
  ungroup() %>%
  select(-size_5)
```

## Portfolios on book-to-market

Form five portfolios on book-to-market within each of the five size portfolios

```{r}
# Compute book_to_market quintiles
book_to_market_quintiles <- data_3 %>%
  filter(!is.na(book_to_market)) %>%
  group_by(size_q, period) %>%
  summarise(
    book_to_market_5 = list(quantile(book_to_market, probs = probs))
  ) %>%
  suppressMessages()

# Join book-to-market quintiles to ticker-period-size_q data
data_4 <- data_3 %>%
  inner_join(book_to_market_quintiles, by = c("period", "size_q"))

# Compute bin in which ticker-size_q-book_to_market combination is put per period
data_5 <- data_4 %>%
  rowwise() %>%
  mutate(
    book_to_market_q = findInterval(
      book_to_market, book_to_market_5, all.inside = T
    )
  ) %>%
  ungroup() %>%
  select(-book_to_market_5)
```

## Portfolios on momentum

```{r}
# Compute momentum quintiles
momentum_quintiles <- data_5 %>%
  filter(!is.na(momentum)) %>%
  group_by(period, size_q, book_to_market_q) %>%
  summarise(momentum_5 = list(quantile(momentum, probs = probs))) %>%
  suppressMessages()

# Join momentum quintiles to ticker-period-size_q data
data_6 <- data_5 %>%
  inner_join(
    momentum_quintiles, 
    by = c("period", "size_q", "book_to_market_q")
  )

# Compute bin in which ticker-size_q-momentum combination is put per period
data_7 <- data_6 %>%
  rowwise() %>%
  mutate(momentum_q = findInterval(momentum, momentum_5, all.inside = T)) %>%
  ungroup() %>%
  select(-momentum_5)
```

# Merge portfolio information with CRSP & trend data

```{r}
ct_2 <- ct %>%
  inner_join(data_7, by = c("ticker", "period"))
```

# Compute portfolio returns

```{r}
ct_3 <- ct_2 %>%
  group_by(period, size_q, book_to_market_q, momentum_q) %>%
  mutate(pf_ret = mean(ret, na.rm = TRUE)) %>%
  ungroup()
```

# Compute abnormal returns

```{r}
ct_4 <- ct_3 %>%
  mutate(abn_ret = ret - pf_ret)
```

# Compute time-aggregated returns

```{r}
# Add column to data containing the number of weeks since Sunday, 1999-12-27
ct_5 <- ct_4 %>%
  mutate(
    n_weeks = difftime(date, as.Date("1999-12-27")) %>% as.numeric %/% 7
  )

ct_6 <- ct_5 %>%
  group_by(ticker, n_weeks) %>%
  mutate(
    weekly_abn_ret = average_return(abn_ret)
  ) %>%
  group_by(ticker, period) %>%
  mutate(
    period_abn_ret = average_return(abn_ret)
  ) %>%
  ungroup()
```


# Save data set

```{r}
write_rds(
  ct_6, 
  file.path(
    "./../large_data", paste_time0(options), "Transform-01-AbnormalReturns.rds"
  )
)
```

